---
title: "Physics Validation of Y3 Clusters Code"
author: "Y3 Cluster Team (3ct)"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: true
      toc_depth: 2
      toc_float: true
      number_sections: true
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      cache = FALSE,
                      fig.align = "center",
                      fig.asp = 1/3)
# library(tidyverse)
library(dplyr)
library(readr)
library(tidyr)
library(stringr)
library(ggplot2)

truecolor = "black"
trueshape = 1
testcolor = "cyan"
testshape = 3
frac.errormargin = 1.0e-6
abs.errormargin = 1.0e-12

testresult <-function(df) {
  npass <- sum(df$pass)
  npoints <- nrow(df)
  str_interp("**For this test, ${sum(df$pass)} of ${nrow(df)} points pass.**")
}

geom_smallpoint <- function(...) geom_point(..., size = 0.5)

nunique <- function(x) length(unique(x))
```

# Introduction

This document presents
demonstrations of the validity of calculations
done in the Y3 clusters analysis code.
Each section contains the validation testing results
for a single function in our calculation.

a. For each of these functions,
we have evaluations from a trusted source that are assumed to be correct.
We test *our implementation* by comparing the results to the evaluations from the trusted source.
a. Unless otherwise indicated, we require fractional differences to be less than
$`r frac.errormargin`$,
**or** absolute differences less than
$`r abs.errormargin`$.
The relative error bound was chosen to be
smaller than the systematic uncertainty
to which the value is known from the trusted source.
An absolute error bound is needed for the cases in which
the predicted value is 0 or very close to 0.
a. In each section, we show both absolute and residual comparisons.
In the absolute comparison plots,
the *expected* (or *true*) values are shown in `r truecolor`,
and results from our *implementation* (or *test* values) are shown in `r testcolor`. We also show the fractional (relative) and absolute residuals.
a. In the plot of the fractional residuals,
the color of each point is determined by whether or not that point
passes the absolute residual test.
In the plot of the absolute residuals,
the color of each point is determined by whether or not that point
passes the fractional residual test.
Thus, in each plot, a point passes the test if *either* it is blue or within
the error bounds (indicated by horizontal lines).
Points that fail the test will be both red and outside of the allowable error bounds (indicated by horizontal lines).
a. For a test to pass, it is required that every calculated point pass.
This condition is verified automatically and the result is noted in bold in each section.


# Part I — Costanzi-2026 pipeline (observable validation)

```{r partI-setup, include=FALSE}
COSTANZI_DIR <- "../validation-data/costanzi2026"

# Source the hand-curated audit / timing tables.  audit_data.R is
# gitignored; if it's missing the audit subsection just shows a stub.
audit_loaded <- tryCatch({ source("audit_data.R"); TRUE },
                        error = function(e) FALSE)

read_observable <- function(name) {
  p <- file.path(COSTANZI_DIR, paste0(name, ".tsv"))
  if (!file.exists(p)) return(NULL)
  readr::read_tsv(p, show_col_types = FALSE)
}

partI_status <- function(df, frac_tol = 0.014, abs_tol = 1e-12) {
  # Costanzi-2026 sections use a looser fractional tolerance (~1.4%
  # for sigma_prj, ~1% for the operators) than the Y1 1e-6 baseline.
  if (is.null(df) || nrow(df) == 0L) {
    return("**Validation TSV not found — run `make data` after fresh `cosmosis cosmosis-models/physics_validation_pipelines.ini` and `cosmosis cosmosis-models/physics_validation_camb.ini` runs.**")
  }
  npts <- nrow(df)
  pass <- abs(df$fracerr) < frac_tol | abs(df$resid) < abs_tol
  npass <- sum(pass, na.rm = TRUE)
  stringr::str_interp("**${npass} of ${npts} points pass at frac < ${frac_tol}.**")
}

partI_residual_plot <- function(df, frac_tol = 0.014,
                                facet_col = "label") {
  if (is.null(df) || nrow(df) == 0L) return(invisible(NULL))
  ggplot(df, aes(x = x_value, y = fracerr,
                 color = factor(.data[[facet_col]]))) +
    geom_smallpoint() +
    geom_hline(yintercept = c(-frac_tol, frac_tol),
               linetype = "dashed", alpha = 0.5) +
    facet_wrap(vars(.data[[facet_col]]), scales = "free_y") +
    labs(x = "wall-grid index", y = "(ytrue - ytest)/ytrue",
         color = facet_col) +
    theme_bw() +
    theme(legend.position = "none")
}

# Two-panel "Part I" plot (Part II's recipe, applied per-observable):
# absolute overlay (black=ytrue, cyan=ytest) + fractional residual.
# `x_col`/`color_col` let each section pick its natural abscissa
# (k_h for cp_camb, M for hmf/halo_bias, ...) and grouping (z bin).
partI_two_panel <- function(df, frac_tol,
                            x_col = "x_value",
                            color_col = "z",
                            x_label = "x",
                            y_label = "y",
                            log_x = TRUE, log_y = TRUE) {
  if (is.null(df) || nrow(df) == 0L) return(invisible(NULL))

  # Order rows by (color group, x) so geom_line connects in the right
  # direction.  Drop non-positive y from the absolute panel when y is
  # being plotted on a log scale.
  df_abs <- df[order(df[[color_col]], df[[x_col]]), ]
  if (log_y) {
    df_abs <- df_abs[df_abs$ytrue > 0 & df_abs$ytest > 0, ]
  }

  p_abs <- ggplot(df_abs, aes(x = .data[[x_col]],
                              color = factor(.data[[color_col]]),
                              group = factor(.data[[color_col]]))) +
    geom_line(aes(y = ytrue), linetype = "solid",  size = 0.6) +
    geom_line(aes(y = ytest), linetype = "dashed", size = 0.6) +
    labs(x = x_label, y = y_label, color = color_col,
         caption = "solid = truth, dashed = test") +
    theme_bw()
  if (log_x) p_abs <- p_abs + scale_x_log10()
  if (log_y) p_abs <- p_abs + scale_y_log10()

  p_res <- ggplot(df, aes(x = .data[[x_col]], y = fracerr,
                          color = factor(.data[[color_col]]),
                          group = factor(.data[[color_col]]))) +
    geom_line(size = 0.6) +
    geom_hline(yintercept = c(-frac_tol, frac_tol),
               linetype = "dashed", alpha = 0.5) +
    labs(x = x_label, y = "(ytrue - ytest)/ytrue",
         color = color_col) +
    theme_bw()
  if (log_x) p_res <- p_res + scale_x_log10()

  print(p_abs); print(p_res)
  invisible(NULL)
}

partI_tbd <- function(why) {
  cat("> *Validation script TBD.* ", why, "\n",
      "> See `emit_validation_tsv.py` stub for the implementation hook.\n",
      sep = "")
}
```

This part validates the **Costanzi-2026 pipeline sourced from
`src/pipelines/`** end-to-end. The numbers behind every subsection come
from the run pair

* `cosmosis-models/physics_validation_pipelines.ini` — cp_camb backend,
  dump `cosmosis-models/physics_validation_pipelines_output/`;
* `cosmosis-models/physics_validation_camb.ini` — CAMB-truth companion
  (same wall, same widePlanck fiducial), dump
  `cosmosis-models/physics_validation_camb_output/`;

in which **every observable stage is a `src/pipelines/` implementation**:
`halo_model` = `src/pipelines/cosmology/halo_model.py`, `prj_params` and
`sel_function` = `src/pipelines/systematics/{selection_function,
selection_richness}`, `bsel` = `src/pipelines/systematics/selection_bias`,
and the observables `NumCountsSijGl`, `NumCounts3d`, `Shear1hGl`,
`ShearPrjGl`, `ShearPrjCuhre` = `src/pipelines/des_y3/{number_counts,
shear_1h2h, shear_projection}`. The allowed non-pipelines inputs (owner
decision, 2026-08-25) are **`MfTinker`** (CSL mass function),
**`cp_camb`** (CosmoPower emulator), and the **survey-area/footprint**
inputs; `consistency`/`GrowthFactor` are CosmoSIS bookkeeping. Two
documented gaps remain sourced from the production tree because no
`src/pipelines` equivalent exists yet: `average_sigma_crit_inv`
(`src/modules/average_sigma_crit_inv/`) and the C++ `b_sel_marg`
operator wall (`src/modules/b_sel_marg_cpu/BSelMargIntegrand.so`) that
`bsel` closes. Both runs build $\xi_{\rm NL}$ from **linear** $P(k)$
(`docs/known_issues/xi_nl_linear_fallback.md`, issue #9): the CAMB
companion deliberately sets `nonlinear = none` so the CAMB-vs-cp_camb
residuals below are pure-emulator checks, not missing-halofit checks.

Algorithms and numerical recipes live in
[`docs/pipeline_modules.tex`](../docs/pipeline_modules.tex); the source-tree
catalog is in [`docs/code_reference.md`](../docs/code_reference.md). Each
subsection compares one pipeline output against an independent reference
(Python `richness_selection`, `cluster_toolkit`, or full CAMB).

Headline budgets (after the six fixes audited in
`pipeline_modules.tex` §12):

```{r partI-headlines, results='asis'}
if (audit_loaded) {
  cat("- $\\Sigma_{\\rm prj} / \\Delta\\Sigma_{\\rm prj}$ agreement:",
      "≤", audit_headlines$sigma_prj_max_pct, "% over",
      "$R \\in [", audit_headlines$R_range_hMpc[1], ", ",
      audit_headlines$R_range_hMpc[2], "]\\,h^{-1}\\,$Mpc\n")
  cat("- $(P_1, I_1, J)$ per-operator: ≤",
      audit_headlines$p_operator_max_pct, "%\n")
  cat("- $(B_{\\rm small}, B_{\\rm large})$ finisher: ~",
      audit_headlines$bsel_finisher_max_pct, "%\n")
  cat("- `cp_camb` $P(k)$: ≤",
      audit_headlines$cp_camb_pk_max_pct, "% across the prior box\n")
  cat("- Per-sample wall: ", audit_headlines$per_sample_baseline_s, "s → ",
      audit_headlines$per_sample_optimum_s, "s ($-",
      audit_headlines$per_sample_speedup_pct, "$%)\n", sep = "")
} else {
  cat("*audit_data.R not found — headline numbers omitted; see `pipeline_modules.tex` §12.*\n")
}
```

## Linear matter power spectrum P(k, z) — `cp_camb`

CosmoPower emulator (`ytest`, read from `cdm_baryon_power_lin/p_k`) vs
full CAMB at the same cosmology (`ytrue`). Comparison is on a 6 z × 25 k
subsample restricted to the cluster lens-redshift range $z \in [0, 1]$.
Headline: ≤0.7% across the prior box (see
[`docs/emulator_validation.tex`](../docs/emulator_validation.tex)).

```{r cp-camb-pk-load, include=FALSE}
df_pk <- read_observable("cp_camb_pk")
```

```{r cp-camb-pk-status, results='asis'}
cat(partI_status(df_pk, frac_tol = 0.007))
```

```{r cp-camb-pk-plot, fig.asp=0.45}
partI_two_panel(df_pk, frac_tol = 0.007,
                x_col = "k_h", color_col = "z",
                x_label = "k  [h/Mpc]",
                y_label = "P_cb(k, z)  [(Mpc/h)^3]",
                log_x = TRUE, log_y = TRUE)
```

## Halo mass function $n(M, z)$ — `mf_tinker`

CosmoSIS `MfTinker` driven by full CAMB (`ytrue`, read from
`physics_validation_camb_output/mass_function/dndlnmh`) vs the same module
driven by the cp_camb emulator (`ytest`, read from
`physics_validation_pipelines_output/mass_function/dndlnmh`). Both use
Tinker 2008 with $\Delta = 200m$ — apples-to-apples through the same
HMF kernel; the only thing changing between truth and test is the
upstream linear $P(k, z)$. Restricted to $M \in [10^{12}, 10^{15}]
\,M_\odot/h$ and $z \le 1.0$.

Headline budget per the camb-emulator/hmf_report (v3c emulator at
widePlanck): max $\sim 2\%$ at high $z$, sub-percent at $z \lesssim
0.5$.

```{r hmf-tinker-load, include=FALSE}
df_hmf <- read_observable("hmf_tinker")
```

```{r hmf-tinker-status, results='asis'}
cat(partI_status(df_hmf, frac_tol = 0.02))
```

```{r hmf-tinker-plot, fig.asp=0.45}
partI_two_panel(df_hmf, frac_tol = 0.02,
                x_col = "M", color_col = "z",
                x_label = "M  [Msun/h]",
                y_label = "dn/dlnM  [(Msun/h)^-1 (cMpc/h)^-3]",
                log_x = TRUE, log_y = TRUE)
```

## Halo bias $b(M, z)$ — `halo_model`

`halomodel/bias` published by `src/pipelines/cosmology/halo_model.py`
(the pipelines halo-model module; verified bit-identical to the
`y3_buzzard/halo_model_cosmosis.py` original on the same datablock)
(`ytest`) vs **`cluster_toolkit.bias.bias_at_M`** (Tinker 2010)
evaluated on the same widePlanck cosmology with the CAMB-driven
`cdm_baryon_power_lin` $P(k, z)$ as input (`ytrue`). Δ = 200m,
$M \in [10^{12}, 10^{15}]\,M_\odot/h$, $z \in [0, 1.0]$. The residual
captures the full pipeline error: emulator $P(k)$, growth-factor
normalisation fix (audit item 2), and CosmoSIS-vs-`cluster_toolkit`
$\sigma(M)$ integrators.

```{r halo-bias-load, include=FALSE}
df_bias <- read_observable("halo_bias")
```

```{r halo-bias-status, results='asis'}
cat(partI_status(df_bias, frac_tol = 0.005))
```

```{r halo-bias-plot, fig.asp=0.45}
partI_two_panel(df_bias, frac_tol = 0.005,
                x_col = "M", color_col = "z",
                x_label = "M  [Msun/h]",
                y_label = "b(M, z)",
                log_x = TRUE, log_y = FALSE)
```

## Number counts $N_i(\lambda, z)$ — `NumCounts3d`

### Pure-emulator check: `NumCounts3d` (CAMB) vs `NumCounts3d` (cp_camb)

Both arms run the **same** brute-force adaptive triple integral
`NumCounts3d` (`src/pipelines/des_y3/number_counts/cpp/3d`, the
pipelines successor of `NumCountsFullScalarIntegrand`) over
$(\ell^t, z, \ln M)$ at the same
widePlanck cosmology; the only difference is the upstream linear
$P(k)$. This mirrors
[`docs/emulator_validation.tex`](../docs/emulator_validation.tex)
Sec. 2 (the v3c emulator validation report), whose Table 1 reports
per-bin residuals 0.23–0.69%.

```{r num-counts-load, include=FALSE}
df_nc <- read_observable("num_counts")
```

```{r num-counts-status, results='asis'}
cat(partI_status(df_nc, frac_tol = 0.01))
```

```{r num-counts-plot, fig.asp=0.45}
partI_two_panel(df_nc, frac_tol = 0.01,
                x_col = "lob", color_col = "label",
                x_label = "lambda (bin centre)",
                y_label = "N_i  [counts]",
                log_x = TRUE, log_y = TRUE)
```

### Production-path check: `NumCounts3d` (CAMB) vs `NumCountsSijGl` (cp_camb)

`ytrue` is the CAMB-driven brute-force triple integral (same as
above), `ytest` is the fast path: cp_camb $\to$ `MfTinker`
$\to$ `sel_function` (closed-form $K_i(\ell^t, z)$ kernel) $\to$
`NumCountsSijGl` (`src/pipelines/des_y3/number_counts/cpp/0d`, the
fixed-GL S_ij-tabulated backend — same algorithm as production
`NumCountsSel` by identity). The residual combines
(i) the cp_camb emulator's $P(k)$ bias (≤1% per the pure-emulator
check above) and (ii) the Sel approximation's faithfulness against
the literal triple integral with full CAMB. After the
`PROJ_Y3_B_i_t` $\to$ `RmSelFunction_t` rewire (May 19) the residual
collapses to **median 0.4%, max 4% (highest-$z$ cells)** — well
inside the cluster-cosmology error budget.

```{r num-counts-sel-load, include=FALSE}
df_ncs <- read_observable("num_counts_sel")
```

```{r num-counts-sel-status, results='asis'}
cat(partI_status(df_ncs, frac_tol = 0.05))
```

```{r num-counts-sel-plot, fig.asp=0.45}
partI_two_panel(df_ncs, frac_tol = 0.05,
                x_col = "lob", color_col = "label",
                x_label = "lambda (bin centre)",
                y_label = "N_i  [counts]",
                log_x = TRUE, log_y = TRUE)
```

## Mass- and bias-weighted $N_i$ — effective mass and bias

```{r mass-bias-weighted, results='asis'}
partI_tbd("MassWeightedSel and BiasWeightedSel against analytic M_eff = ⟨M⟩_S, b_eff = ⟨b⟩_S on the same 4×3 grid (no src/pipelines backend exists yet; modules would also need adding to physics_validation_pipelines.ini).")
```

## $(\mathcal{P}[1], I_1, J)$ — selection-bias operators from `b_sel_marg`

The three P[X] operator walls published by the C++ `b_sel_marg` module
(`BSelMargIntegrand.so`, sections `b_sel_marg_{p1, i1, j}`;
$J \equiv I_2 - I_1$ computed directly to avoid cancellation), compared
against the Python `richness_selection.SelBias._P_operator` reference
evaluated on the same datablock inputs (same HMF, halo bias, $\xi_{\rm
NL}$, and the dump's own `cluster_mor` HOD point). 12 wall rows
$\times$ 3 operators = 36 cells.

```{r p-operator-load, include=FALSE}
df_pop <- read_observable("p_operator")
```

```{r p-operator-status, results='asis'}
cat(partI_status(df_pop, frac_tol = 0.05))
```

```{r p-operator-plot, fig.asp=0.45}
partI_residual_plot(df_pop, frac_tol = 0.05, facet_col = "label")
```

## $\overline{b}^{\rm LOS}_{ij}, \overline{b}^{\rm LSS}_{ij}$ — sigmoid plateaus from `bsel`

The two latent-marginalised plateau amplitudes of the
mixture-of-regimes ansatz [`projection_lensing_paper.tex`
Eq.~b_sigmoid, Eq.~b_marg_lt] are published by the Python `bsel`
finisher (`src/pipelines/systematics/selection_bias/python/bsel.py`)
under the datablock keys
`b_sel_marginalised/{b_small, b_large}`, with the mapping

* `b_small` $\leftrightarrow \overline{b}^{\rm LOS}_{ij}$ (small-$\theta$ /
  correlated-LOS asymptote),
* `b_large` $\leftrightarrow \overline{b}^{\rm LSS}_{ij}$ (large-$\theta$ /
  LSS asymptote).

This subsection is an **$\ell^t$-integration precision test of the
finisher**. Both `ytrue` and `ytest` consume the *same* upstream C++
operators $(\mathcal{P}[1], I_1, J \equiv I_2 - I_1)$ from §2.5 and the
*same* `bias_from_precomp` per-$\ell^t$ plateau closures
$\overline{b}^{\rm LSS}(\lambda^o,\ell^t,z^o)$ /
$\overline{b}^{\rm LOS}(\lambda^o,\ell^t,z^o)$
[paper Eqs.~b_LSS_latent, b_LOS_latent]; they differ only in the
$\ell^t$-marginalisation quadrature against the bin posterior
$P(\ell^t|\Delta\lambda_i,\Delta z_j) \propto
\mathcal{S}_{\Delta\lambda_i}(\ell^t,z^o)\,\pi(\ell^t,z^o)$ with
$\pi(\ell^t,z^o) = \int dM\,n(M,z^o)\,P(\ell^t|M,z^o)$
[paper Eq.~plt_post]:

* `ytest` — production path: per-$\lambda^o$ Gauss–Legendre on
  $\ell^t \in [1,\,3\lambda^o]$.
* `ytrue` — `scipy.integrate.quad` (`epsabs=1e-12, epsrel=1e-9`) over
  the same integrand on the same support.

Grid: 4 $\lambda^o$-bins $\times$ 3 $z^o$-bins $\times$ 2 plateaus =
24 cells. The full C++/Python equivalence test at the assembled
$\overline{b}_{{\rm sel},ij}(\theta)$ is §2.8. Pointer:
`pipeline_modules.tex` §9.

```{r bsel-plateaus-load, include=FALSE}
df_plat <- read_observable("bsel_plateaus")
```

```{r bsel-plateaus-status, results='asis'}
cat(partI_status(df_plat, frac_tol = 0.05))
```

```{r bsel-plateaus-plot, fig.asp=0.45}
partI_two_panel(df_plat, frac_tol = 0.05,
                x_col = "lob", color_col = "label",
                x_label = "lambda^o (bin centre)",
                y_label = "B (sigmoid plateau)",
                log_x = TRUE, log_y = FALSE)
```

## $b_{\rm sel,marg}(\theta)$ closure — `bsel`

**Retired with the theta table.** The current `b_sel_marg` +
`bsel` pair no longer publishes a `b_sel_marginalised/{theta, vals}`
grid: the $\theta$ dependence factorizes exactly in the C++ operators,
so `bsel` publishes only the two plateau scalars
`b_sel_marginalised/{b_small, b_large}` per wall row and every consumer
(ShearPrjGl, ShearPrjCuhre) assembles
$b_{\rm sel,marg}(\theta) = \overline{B_0}\,(1 - \sigma(\theta)) +
\overline{B_\infty}\,\sigma(\theta)$ analytically in-integrand. The
plateau amplitudes themselves are validated in the previous section;
there is no longer a published $\theta$-grid to pin. (The historical
audit-item-7 record below is kept for provenance.)

> **Audit item 7 (fixed 2026-05-19).** `y3_buzzard/bsel.py` previously
> read `distances/d_c` directly (Mpc) when computing
> $\theta_\lambda = R_\lambda(\lambda)\,(1+z)/\chi(z)$ for the
> $\sigma(\theta)$ tabulation, while the C++ consumers
> (`p_operator_t.hh`, `sigma_prj_t.hh`) explicitly multiply
> `distances/d_c` by $h_0$ to put $\chi$ in cMpc$/h$.
> The Python tabulation therefore used $\theta_\lambda$ smaller by
> $h_0 = 0.6766$, shifting the sigmoid transition relative to where the
> C++ side evaluates it. Residual was median **12%, max 37%**
> pre-fix; collapses to median **3.2%, max 3.5%** here (32/32 within
> 5%, matching the §2.6 plateau-only residual).

```{r bsel-load, include=FALSE}
df_bsel <- read_observable("bsel_theta")
```

```{r bsel-status, results='asis'}
cat(partI_status(df_bsel, frac_tol = 0.05))
```

```{r bsel-plot, fig.asp=0.45}
partI_two_panel(df_bsel, frac_tol = 0.05,
                x_col = "theta_over_lam", color_col = "label",
                x_label = "theta / theta_lambda",
                y_label = "b_sel,marg(theta)",
                log_x = TRUE, log_y = FALSE)
```

## $\Sigma^{1h}(R)$ — 1-halo lensing surface density

**Not emitted by the pipelines backend.** The des_y3 1-halo module
`Shear1hGl` (`src/pipelines/des_y3/shear_1h2h/cpp/0d`) publishes the
observable $\gamma_t^{1h}$ only — there is no pipelines producer of a
standalone $\Sigma^{1h}$ wall (production `Sigma1hSel` is not part of
the physics_validation pipeline). The 1-halo physics is covered by the
$\gamma_t^{1h}$ section below and, at the profile-table level, by
`test/halo_model.test.py` / `test/shear1h_gl.test.cc`.

```{r sigma-1h-load, include=FALSE}
df_s1h <- read_observable("sigma_1h")
```

```{r sigma-1h-status, results='asis'}
cat(partI_status(df_s1h, frac_tol = 0.01))
```

```{r sigma-1h-plot, fig.asp=0.45}
partI_two_panel(df_s1h, frac_tol = 0.01,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "Sigma_1h(R)",
                log_x = TRUE, log_y = TRUE)
```

## $\Delta\Sigma^{1h}(R)$ — 1-halo $\Delta\Sigma$

**Not emitted by the pipelines backend** — same status as the
$\Sigma^{1h}$ section above (`Shear1hGl` publishes $\gamma_t^{1h}$
only).

```{r dsigma-1h-load, include=FALSE}
df_ds1h <- read_observable("dsigma_1h")
```

```{r dsigma-1h-status, results='asis'}
cat(partI_status(df_ds1h, frac_tol = 0.01))
```

```{r dsigma-1h-plot, fig.asp=0.45}
partI_two_panel(df_ds1h, frac_tol = 0.01,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "Delta Sigma_1h(R)",
                log_x = TRUE, log_y = TRUE)
```

## $\gamma_t^{1h}(R)$ — 1-halo tangential shear

CAMB-driven `Shear1hGl` (`ytrue`) vs cp_camb-driven `Shear1hGl`
(`ytest`) — the fixed-GL des_y3 backend
(`src/pipelines/des_y3/shear_1h2h/cpp/0d`, bitwise-identical algorithm
to production `Shear1hMisSel`), with the DES-Y3 gamma-kernel
miscentering mixture from the `[miscentering]` values. Tolerance 5%
absorbs the noise floor at the largest-$R$ bin where $\gamma_t^{1h}$
is tiny (C++ rounding limit).

```{r shear-1h-load, include=FALSE}
df_sh1h <- read_observable("shear_1h")
```

```{r shear-1h-status, results='asis'}
cat(partI_status(df_sh1h, frac_tol = 0.05))
```

```{r shear-1h-plot, fig.asp=0.45}
partI_two_panel(df_sh1h, frac_tol = 0.05,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "gamma_t^1h(R)",
                log_x = TRUE, log_y = TRUE)
```

## $\Sigma_{\rm prj}$ — two-halo projection

### Pure-emulator check: `ShearPrjCuhre` (CAMB) vs (cp_camb)

CAMB-driven and cp_camb-driven `sigma_prj_cuhre` outputs
(`ShearPrjCuhre`, `src/pipelines/des_y3/shear_projection/cpp/2d` — the
only $\Sigma_{\rm prj}$ producer in the pipelines run; the fixed-GL
`ShearPrjGl` publishes $\Delta\Sigma_{\rm prj}$ and
$\gamma_t^{\rm prj}$ only). Same backend on both arms, only $P(k)$
differs. Headline: $\le 1.4\%$ over $R \in [0.1, 30]\,h^{-1}$ Mpc per
`pipeline_modules.tex` §10.

```{r sigma-prj-load, include=FALSE}
df_sp <- read_observable("sigma_prj")
```

```{r sigma-prj-status, results='asis'}
cat(partI_status(df_sp, frac_tol = 0.05))
```

```{r sigma-prj-plot, fig.asp=0.5}
partI_two_panel(df_sp, frac_tol = 0.05,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "Sigma_prj(R)",
                log_x = TRUE, log_y = TRUE)
```

### Integrator-precision check: `ShearPrjCuhre` (cp_camb) vs `ShearPrjGl` (cp_camb)

**Not available for $\Sigma_{\rm prj}$:** the fixed-GL `ShearPrjGl`
does not publish a $\Sigma_{\rm prj}$ wall, so there is no evaluator
to compare `sigma_prj_cuhre` against. The Cuhre-vs-GL check runs on
the $\Delta\Sigma_{\rm prj}$ and $\gamma_t^{\rm prj}$ sections
below — see the normalisation-drift flag there.

```{r sigma-prj-cuhre-load, include=FALSE}
df_spc <- read_observable("sigma_prj_cuhre")
```

```{r sigma-prj-cuhre-status, results='asis'}
cat(partI_status(df_spc, frac_tol = 0.05))
```

```{r sigma-prj-cuhre-plot, fig.asp=0.5}
partI_two_panel(df_spc, frac_tol = 0.05,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "Sigma_prj(R)",
                log_x = TRUE, log_y = TRUE)
```

## $\Delta\Sigma_{\rm prj}$ — two-halo $\Delta\Sigma$

### Pure-emulator check: `ShearPrjGl` (CAMB) vs (cp_camb)

CAMB-driven vs cp_camb-driven `dsigma_prj_gl` (`ShearPrjGl`,
`src/pipelines/des_y3/shear_projection/cpp/0d` — same
`sp_detail::ShearPrjCore` as production `shear_prj_frozen_physics`).
With the CAMB companion pinned to linear $\xi_{\rm NL}$ (issue #9
apples-to-apples, see the Part I preamble) the residual is the pure
emulator effect: **median 0.02%, max 0.07%** on this run pair.

> **Audit item 9 (historical) — bsel(θ) residual amplification.** $\Delta\Sigma_{\rm prj}$
> shows a 14% max residual concentrated at small $R$ in low-$\lambda$ bins
> while $\Sigma_{\rm prj}$ (§2.10) sits at 3.5% max. Decomposition: the
> `rnd` piece (uniform-foreground, no $b_{\rm sel}$ factor) agrees to
> **0.2%**; the `cl` piece (clustering excess, multiplied by
> $b_{\rm sel}(\theta)$) carries the entire residual at **median 5.5%,
> max 19.5%**. This is the §2.7 $b_{\rm sel}(\theta)$ residual ceiling
> (~3% even after the chi-unit fix in audit item 7) propagated through
> the small-$R$ NFW-$\Delta\Sigma$ kernel that weights the inner-$\theta$
> part of the sigmoid most heavily. Closing the production budget
> ($\le 1.4\%$, `pipeline_modules.tex` §10) requires tightening the
> `bsel.py` plateau marginalisation, not the emulator $P(k)$.

```{r dsigma-prj-load, include=FALSE}
df_dsp <- read_observable("dsigma_prj")
```

```{r dsigma-prj-status, results='asis'}
cat(partI_status(df_dsp, frac_tol = 0.05))
```

```{r dsigma-prj-plot, fig.asp=0.5}
partI_two_panel(df_dsp, frac_tol = 0.05,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "Delta Sigma_prj(R)",
                log_x = TRUE, log_y = TRUE)
```

### Integrator-precision check: `ShearPrjCuhre` vs `ShearPrjGl` (cp_camb)

> **🚨 FLAGGED — `ShearPrjCuhre` normalisation drift (issue #23,
> backend status "pending" in
> `src/pipelines/des_y3/shear_projection/README.md`).** The Cuhre
> adaptive backend and the fixed-GL `ShearPrjGl` should agree to the
> Cuhre tolerance ($\sim 10^{-3}$) on the same datablock. They instead
> disagree by a flat factor $\approx 0.46$ (GL $\approx 2.15 \times$
> Cuhre) across **all** 120 cells (median
> $|\Delta|/{\rm GL} \approx 1.17$), `cl` and `rnd` scaled equally —
> a multiplicative normalisation on the Cuhre code path, unchanged
> from the pre-pipelines smoke runs. Treat as a known offset until the
> 2d backend's pending re-validation lands.

```{r dsigma-prj-cuhre-load, include=FALSE}
df_dspc <- read_observable("dsigma_prj_cuhre")
```

```{r dsigma-prj-cuhre-status, results='asis'}
cat(partI_status(df_dspc, frac_tol = 0.05))
```

```{r dsigma-prj-cuhre-plot, fig.asp=0.5}
partI_two_panel(df_dspc, frac_tol = 0.05,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "Delta Sigma_prj(R)",
                log_x = TRUE, log_y = TRUE)
```

## $\gamma_t^{\rm prj}$ — two-halo tangential shear

### Pure-emulator check: `ShearPrjGl` (CAMB) vs (cp_camb)

$\gamma_t^{\rm prj}(R) = \Delta\Sigma_{\rm prj}(R)\,\Sigma_{\rm crit}^{-1}$
with $\Sigma_{\rm crit}^{-1}$ computed from the source $p(z)$, identical
across both arms — so the **fractional residual is exactly the
$\Delta\Sigma_{\rm prj}$ residual** (cell-for-cell, see §2.11). The
absolute values are $\sim 10^{-18}$ to $10^{-16}$ (after the $\Sigma_{\rm
crit}^{-1}$ rescale), so most cells pass the absolute-tolerance leg of
the test ($|{\rm resid}| < 10^{-12}$) regardless of fractional error.
Read the §2.11 fractional residual instead for the physics signal.

```{r shear-prj-load, include=FALSE}
df_shp <- read_observable("shear_prj")
```

```{r shear-prj-status, results='asis'}
cat(partI_status(df_shp, frac_tol = 0.05))
```

```{r shear-prj-plot, fig.asp=0.5}
partI_two_panel(df_shp, frac_tol = 0.05,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "gamma_t^prj(R)",
                log_x = TRUE, log_y = TRUE)
```

### Integrator-precision check: `ShearPrjCuhre` vs `ShearPrjGl` (cp_camb)

Inherits the same $\approx 0.46$ normalisation offset flagged in the
$\Delta\Sigma_{\rm prj}$ Cuhre subsection (issue #23).

```{r shear-prj-cuhre-load, include=FALSE}
df_shpc <- read_observable("shear_prj_cuhre")
```

```{r shear-prj-cuhre-status, results='asis'}
cat(partI_status(df_shpc, frac_tol = 0.05))
```

```{r shear-prj-cuhre-plot, fig.asp=0.5}
partI_two_panel(df_shpc, frac_tol = 0.05,
                x_col = "R", color_col = "label",
                x_label = "R [cMpc/h]",
                y_label = "gamma_t^prj(R)",
                log_x = TRUE, log_y = TRUE)
```

Pointer: `pipeline_modules.tex` §10 (line 742); the per-$R$ `rnd`/`cl`
decomposition is in `docs/sigma_prj_py_vs_cpp_report.tex`.

## Pipeline timing & precision audit

```{r audit-table, results='asis'}
if (audit_loaded) {
  cat("**Six-bug debugging arc** (transcribed from `pipeline_modules.tex` §12.3):\n\n")
  knitr::kable(audit_bugs, format = "markdown")
  cat("\n\n**Per-module wall-time trajectory** (§12.1):\n\n")
  knitr::kable(audit_timing_trajectory, format = "markdown",
               digits = 3)
  cat("\n\n**`n_lnm` convergence sweep** (§12.2 — the whole-pipeline ",
      "optimum at $n_{\\ln M} = ", audit_headlines$n_lnm_optimum,
      "$; $n_{\\ln M} = ", audit_headlines$n_lnm_resonance_avoid,
      "$ hits a GL resonance and must be avoided):\n\n", sep = "")
  knitr::kable(n_lnm_sweep, format = "markdown", digits = 3)
} else {
  cat("*audit_data.R not present in this checkout — see `docs/pipeline_modules.tex` §12 for the timing & precision audit.*\n")
}
```

# Part II — Per-term Y1 baseline (appendix)

The sections below validate the *underlying terms* used inside the
Costanzi-2026 operators (`mor_t2`, `lc_lt_t`, `lo_lc_t`,
`sigma_crit_inverse_t`, etc.). They are kept as the term-level
regression suite — Part I is the headline end-to-end view.

## Survey Area

### SDSS Survey footprint function

```{r}
local.frac.errormargin = frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r, include = FALSE}
omega_z <- read_tsv("../validation-data/omega_z_sdss.out")
omega_z <- mutate(omega_z,
                  resid = ytrue - ytest,
                  fracerr = resid/ytrue,
                  lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                  hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                  within.frac = (abs(fracerr) < local.frac.errormargin),
                  within.abs = (abs(resid) < local.abs.errormargin),
                  pass = within.frac | within.abs)
```

$\Omega_z$ is the area of the survey as a function of redshift $z$.
This section tests our implementation of the SDSS survey area function, which is implemented in the file `omega_z_sdss.hh`.

For this testing, we compare our function output to the independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019).

a. The comparisons are done at `r nrow(omega_z)` unique $z$ points. 
a. Over the range tested
$(`r min(omega_z$z)` < z < `r max(omega_z$z)`)$,
the maximum magnitude of the absolute residual between the expected value
and our approximation is
$`r max(abs(omega_z$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(omega_z$fracerr))`$.
a. For this test, we require the relative residuals to be less than
$`r local.frac.errormargin`$,
or the absolute residuals to be less than
$`r local.abs.errormargin`$.
a. `r testresult(omega_z)`

```{r fig.cap = "Absolute comparison of our implementation results (cyan points) to the expected values (black points) for $\\Omega_z$. Note that over-plotting may obscure the expected values."}
ggplot(omega_z) +
  geom_smallpoint(aes(x = z, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = z, y = ytest), color = testcolor, shape = testshape) +
  xlab(expression(z)) +
  ylab(expression(Omega))
```

```{r fig.cap = "Relative residual between our implementation results and the expected values. Blue horizontal lines are the error tolerance bounds. Blue points pass, and red points fail, the absolute residual test."}
ggplot(omega_z, aes(x = z, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab(expression(z)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

## Halo Mass Function



```{r}
local.frac.errormargin = 0.01
local.abs.errormargin = 1.0e-16
```

```{r include = FALSE}
hmf <- read_tsv("../validation-data/masses_test.out")
hmf$zbin <- as.factor(hmf$z)
hmf <- mutate(hmf,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$n(M_{200m}, z)$ is the halo mass function, which describes the 3D cosmic density of halos at a given mass $M_{200m}$ and $z$. 
We have modified the CosmoSIS halo mass function module to consider the effect of massive neutrinos (Costanzi et al. 2013)- we use the $P_{cdm}$ power spectrum rather than the $P_{tot}$.

For this testing, we compare results from our implementation to the  independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019)

a. The comparisons are made at `r nunique(hmf$z)` unique $z$ values with `r nlevels(hmf$zbin)` different mass values.
a. Over the range tested 
($`r min(hmf$mass)` < M < `r max(hmf$mass)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(hmf$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(hmf$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(hmf)`

Each panel in the figure below corresponds to a different $z$ value, specified in the panel top label.

```{r fig.asp=0.4,fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(hmf) +
  geom_smallpoint(aes(x = mass, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = mass, y = ytest), color = testcolor, shape = testshape) +
  xlab(expression(mass)) +
  ylab("Number Density") +
  scale_y_log10() +
  facet_wrap(zbin~ .)
```

```{r fig.asp=0.4,fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(hmf, aes(x = mass, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(M)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(zbin~ .)
```

**This test requires running masses_test.ini in cosmosis_tests first**
Need to do better than this in the future.

## Volume Element

```{r}
local.frac.errormargin = 1.0e-3
local.abs.errormargin = abs.errormargin
```

```{r, include = FALSE}
dvdodz <- read_tsv("../validation-data/dvdodz_test.out")
dvdodz <- mutate(dvdodz,
                  resid = ytrue - ytest,
                  fracerr = resid/ytrue,
                  lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                  hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                  within.frac = (abs(fracerr) < local.frac.errormargin),
                  within.abs = (abs(resid) < local.abs.errormargin),
                  pass = within.frac | within.abs)
```

$\frac{dV}{d\Omega dz}$ is the co-moving volume element in solid angle $d \Omega$ and redshift interval $dz$, and is a function of $z$. Our implementation is in the file `dv_do_dz_t.hh`.

For this testing, we compare our function output to the independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019), calculated with $\Omega_m =0.3$,  $\Omega_\Lambda =0.7$, $h_0=0.7$.

a. The comparisons are done at `r nrow(dvdodz)` unique $z$ values.
a. Over the range tested
$(`r min(dvdodz$z)` < z < `r max(dvdodz$z)`)$,
the maximum magnitude of the absolute residual between the expected value
and our approximation is
$`r max(abs(dvdodz$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(dvdodz$fracerr))`$.
a. For this test, we require the relative residuals to be less than
$`r local.frac.errormargin`$,
or the absolute residuals to be less than
$`r local.abs.errormargin`$.
a. `r testresult(dvdodz)`

```{r fig.cap = "Absolute comparison of our approximation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(dvdodz) +
  geom_smallpoint(aes(x = z, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = z, y = ytest), color = testcolor, shape = testshape) +
  xlab(expression(z)) +
  ylab(expression(dvdodz))
```

```{r fig.cap = "Relative residual comparison or implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(dvdodz, aes(x = z, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  ylim(-local.frac.errormargin, local.frac.errormargin) +
  xlab(expression(z)) +
  ylab("fractional residual") +
  geom_hline(yintercept = -local.frac.errormargin, color="blue") +
  geom_hline(yintercept =local.frac.errormargin, color="blue")
```

## Scaling Relations

### Mass-observable relation

```{r}
local.frac.errormargin = 5 * frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
mor_t2 <- read_tsv("../validation-data/mor_tt_test.out")
mor_t2$mbin <- as.factor(mor_t2$log10m/log(10))
mor_t2 <- mutate(mor_t2,
                 resid = probtrue - probtest,
                 fracerr = resid/probtrue,
                 lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$P(\lambda |M_{200m})$ describes the probabilistic distribution of the cluster observable richness ($\lambda$) at a given mass $M_200m$. 
Our calculation is implemented in the file `mor_t2.hh`.

For this testing, we compare results from our implementation to the  independent function evaluations from Matteo Costanzi in the SDSS cluster cosmology analysis (Costanzi et al. 2019)

a. The comparisons are made at `r nunique(mor_t2$lambda)` unique $\lambda$ values with `r nlevels(mor_t2$mbin)` different mass values.
a. Over the range tested 
($`r min(mor_t2$lambda)` < \lambda < `r max(mor_t2$lambda)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(mor_t2$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(mor_t2$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(mor_t2)`

Each panel in the figure below corresponds to a different $\mathrm{log} M_{200m}$ value, specified in the panel top label.

```{r fig.asp=0.4,fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(mor_t2) +
  geom_smallpoint(aes(x = lambda, y = probtrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = lambda, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(lambda)) +
  ylab("probability") +
  facet_wrap(mbin~.)
```

```{r fig.asp=0.4,fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(mor_t2, aes(x = lambda, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(lambda)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(mbin~ .)
```

```{r fig.asp=0.4, fig.cap = "Absolute residual comparison of our approximation to the expected results for MOR. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(mor_t2, aes(x= lambda, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab(expression(lambda)) +
  facet_wrap(mbin ~ .)
```

### Projection effect on Scaling relations

```{r}
local.frac.errormargin =  frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
lc_lt <- read_tsv("../validation-data/lc_lt_test.out")
lc_lt$lbin <- as.factor(lc_lt$lambda_true)
lc_lt <- mutate(lc_lt,
                 resid = probtrue - probtest,
                 fracerr = resid/probtrue,
                 lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

Projections of LOS structures cause a shift and an extra scatter to the values of cluster richness $\lambda$. $P(\lambda_c |\lambda_t)$ describes the probabilistic distribution of the richness scatter (between $\lambda_c$ and $\lambda_t$) caused by projections. 
Our calculation is implemented in the file `lc_lt_t.hh`.

For this testing, we compare results from our implementation to the  independent function evaluations from Matteo Costanzi's projection analyses (Costanzi et al. 2019)

a. The comparisons are made at `r nunique(lc_lt$lambda_cen)` unique $\lambda_{c}$ values with `r nlevels(lc_lt$lbin)` different $\lambda_{t}$ values.
a. Over the range tested 
($`r min(lc_lt$lambda_cen)` < \lambda_{cen} < `r max(lc_lt$lambda_cen)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(lc_lt$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(lc_lt$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
 $`r local.abs.errormargin`$.
a. `r testresult(lc_lt)`

Each panel in the figure below corresponds to a $\lambda_{t}$ value,  specified in the panel top label.

```{r fig.asp=0.7, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(lc_lt) +
  geom_smallpoint(aes(x = lambda_cen, y = probtrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = lambda_cen, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(lambda_c)) +
  ylab("probability") +
  facet_wrap(lbin~.)
```

```{r fig.asp=0.7, fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(lc_lt, aes(x = lambda_cen, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(lambda_c)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(lbin~ .)
```

```{r fig.asp=0.7, fig.cap = "Absolute residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(lc_lt, aes(x= lambda_cen, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab(expression(lambda_c)) +
  facet_wrap(lbin ~ .)
```

### Mis-centering effect on scaling relations

```{r}
local.frac.errormargin =  frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
lo_lc <- read_tsv("../validation-data/lo_lc_test.out")
lo_lc$Rbin <- as.factor(lo_lc$Rmis)
lo_lc <- mutate(lo_lc,
                 resid = probtrue - probtest,
                 fracerr = resid/probtrue,
                 lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

Mis-centering cause a shift and an extra scatter to the values of cluster richness $\lambda$. $P(\lambda_o/\lambda_c | R_{mis})$ describes the probabilistic distribution of the richness shift ($\lambda_o/\lambda_c$) with a mis-centering distance $R_{mis}$ 
Our calculation is implemented in the file `lo_lc_t.hh`.

For this testing, we compare results from our implementation to the  independent function evaluations from Yuanyuan Zhang's mis-centering analysis (Zhang et al. 2019).

a. The comparisons are made at `r nunique(lo_lc$deltal)` unique values of $\lambda_{o}/\lambda_{c}$ with `r nlevels(lo_lc$Rbin)` different $R_{mis}$ values.
a. Over the range tested 
($`r min(lo_lc$deltal)` < \lambda_{cen} < `r max(lo_lc$deltal)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(lo_lc$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(lo_lc$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(lo_lc)`

Each panel in the figure below corresponds to a $R_{mis}$ value, specified in the panel top label.

```{r fig.asp=0.7, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(lo_lc) +
  geom_smallpoint(aes(x = deltal , y = probtrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = deltal , y = probtest), color = testcolor, shape = testshape) +
  scale_y_log10() +
  scale_x_log10() +
  xlab(expression(lambda_miscent/lambda_cent)) +
  ylab("probability") +
  facet_wrap(Rbin~.)
```

```{r fig.asp=0.7, fig.cap = "Relative residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(lo_lc, aes(x = deltal, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  xlab(expression(lambda_miscent/lambda_cent)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(Rbin~ .)
```

```{r fig.asp = 0.7, fig.cap = "Absolute residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(lo_lc, aes(x= deltal, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab(expression(lambda_miscent/lambda_cent)) +
  facet_wrap(Rbin ~ .)
```

## Number Counts and Masses in SDSS

**This test requires running sdss_analysis.ini in cosmosis_tests first**.

```{r}
local.frac.errormargin =  0.05
local.abs.errormargin = abs.errormargin
```

In this section, we compare the stacked number counts and mass predictions in stacked richness/mass bins for a SDSS like analysis. These values are the integrated results using components listed above. For this testing, we compare results from our implementation to the implementations from Matteo Costanzi for the SDSS cosmology analysis (Costanzi et al. 2019). The test is conducted for the following set of cosmology and scaling relation parameters specified in **cosmosis_tests/values_comp.ini**.

```{r include = FALSE}
inte <- read_tsv("../validation-data/integrated_bin_test.out")
inte <- mutate(inte,
               lambda = (llo+lhi)/2.0,
               residn = yntrue - yntest,
                 fracerrn = residn/yntrue,
                 residm = ymtrue - ymtest,
                 fracerrm = residm/ymtrue,
                 lo = pmin(-local.frac.errormargin * yntrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * yntrue, local.abs.errormargin),
                 within.frac = (abs(fracerrn) < local.frac.errormargin),
                 within.abs = (abs(residn) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

a. The comparisons are made at `r nunique(inte$lambda)` unique values of $\lambda$ bins.
a. Over the **number of bins** tested and for the stacked **number counts**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residn))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrn))`$.
a. Over the **number of bins** tested and for the stacked **average masses**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residm))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrm))`$. 

```{r fig.asp=0.25, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = yntrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = yntest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab('Richness') +
  ylab("Number Counts")
```

```{r fig.cap = "Relative residual implementation or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrn, color = within.abs)) +
  geom_point() +
  guides(color = "none") +
  scale_x_log10() +
  xlab("richness") +
  ylab("fractional residual for number counts") +
  ylim(-5 * local.frac.errormargin, 5 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

```{r fig.asp=0.25, fig.cap = "Absolute comparison of our implementation to the expected results for average masses. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = ymtrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = ymtest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  xlab('Richness') +
  ylab("LogMass")
```

```{r fig.cap = "Relative residual implementation or our approximation to the expected results for average masses. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrm, color = within.abs)) +
  geom_point() +
  scale_x_log10() +
  guides(color = "none") +
  xlab("richness") +
  ylab("fractional residual for average mass") +
   ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

## Number Counts and Masses in DESY1

**This test requires running y1_analysis.ini in cosmosis_tests first**

```{r}
local.frac.errormargin =  0.05
local.abs.errormargin = abs.errormargin
```

In this section, we compare the stacked number counts and mass predictions in stacked richness/mass bins for a y1 like analysis. These values are the integrated results using components listed above. For this testing, we compare results from our implementation to the implementations from Matteo Costanzi for the y1 cosmology analysis (DES 2019). The test is conducted for the following set of cosmology and scaling relation parameters specified in **cosmosis_tests/values_comp_y1.in**

```{r include = FALSE}
inte <- read_tsv("../validation-data/integrated_bin_y1_test.out")
inte$zbin <- as.factor(inte$zlo)
inte <- mutate(inte,
               lambda = (llo+lhi)/2.0,
               residn = yntrue - yntest,
                 fracerrn = residn/yntrue,
                 residm = ymtrue - ymtest,
                 fracerrm = residm/ymtrue,
                 lo = pmin(-local.frac.errormargin * yntrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * yntrue, local.abs.errormargin),
                 within.frac = (abs(fracerrn) < local.frac.errormargin),
                 within.abs = (abs(residn) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

a. The comparisons are made at `r nunique(inte$lambda)` unique values of $\lambda$ bins and `r nunique(inte$zlo)` unique values of $z$ bins.
a. Over the **number of bins** tested and for the stacked **number counts**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residn))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrn))`$.
a. Over the **number of bins** tested and for the stacked **average masses**,
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(inte$residm))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(inte$fracerrm))`$. 

```{r fig.asp=0.3, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = yntrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = yntest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  xlab('Richness') +
  ylab("Number Counts")+
  facet_wrap(zbin~.)
```

```{r fig.asp=0.3, fig.cap = "Relative residual implementation or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrn, color = within.abs)) +
  geom_point() +
  scale_x_log10() +
  guides(color = "none") +
  xlab("richness") +
  ylab("fractional residual for number counts") +
   ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")+
  facet_wrap(zbin~.)
```

```{r fig.asp=0.3, fig.cap = "Absolute comparison of our implementation to the expected results for average masses. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(inte) +
  geom_point(aes(x =lambda , y = ymtrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = lambda , y = ymtest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  xlab('Richness') +
  ylab("LogMass")+
  facet_wrap(zbin~.)
```

```{r fig.asp=0.3, fig.cap = "Relative residual implementation or our approximation to the expected results for average masses. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(inte, aes(x = lambda, y = fracerrm, color = within.abs)) +
  geom_point() +
  scale_x_log10() +
  guides(color = "none") +
  xlab("richness") +
  ylab("fractional residual for average mass") +
   ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")+
  facet_wrap(zbin~.)
```

## Cluster Mass Profile

The Cluster Mass profiles are calculated by a **separate CosmoSIS module**, which is a python wrapper to Tom McClintock's Cluster_toolkit code. The code exists in the folder `delta_sigma`.

**Before your run this test, make sure to go to the folder `deltasigma` and run
`cosmosis deltasigma_test.ini` to produce the proper test files.**
If you are using the supplied `GNUmakefile` to build the document,
this is done automatically.

### $\xi$ NFW

```{r}
local.frac.errormargin = 0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
xi_nfw <- read_tsv("../validation-data/deltasigma_xi_nfwonly.out")
print(xi_nfw)
xi_nfw <- mutate(xi_nfw,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\xi_{NFW}(r)$ is the cluster matter correlation function in the 1-halo regime, adopting an NFW model. 

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package, for a cluster of $M_{200m}=3.199267137797384375e+14 M_\odot/h$ and $c=5$.

a.  The comparisons are made at `r nrow(xi_nfw)` unique values of $r$. 
a. Over the range tested 
($`r min(xi_nfw$r)` < r < `r max(xi_nfw$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(xi_nfw$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(xi_nfw$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(xi_nfw)`

The values of $\xi$ cover a very wide range, so a logarithmic $y$-axis is needed.
However, many points in the plot have $\xi = 0$.
In order to allow the use of a logarithmic axis, these points are suppressed.

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close. Note that we suppress negative values of $\\xi$ in this plot, so that we can use a log scale."}
ggplot(filter(xi_nfw, ytrue > 0.0, ytest > 0.0)) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab(expression(r)) +
  ylab(expression(xi))
```

```{r fig.cap = "Relative residual implementation or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(filter(xi_nfw, ytrue > 0.0, ytest > 0.0),
      aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  xlab(expression(r)) +
  ylab("fractional residual") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

### $\xi$ 2-halo

```{r}
local.frac.errormargin = 0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
xi_2h <- read_tsv("../validation-data/deltasigma_xi_mm_2halo.out")
# We limit the range of R in our plots to be only positive values of xi_mm. 
# In the generation of the xi_mm_2halo.out file, we cut at < 125 Mpc, where the fiducial xi_mm goes negative
print(xi_2h)
xi_2h <- mutate(xi_2h,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\xi_{mm}(r)$ is the matter correlation function in the cluster 2-halo regime, adopting a non-linear matter power spectrum, and not accounting for cluster bias.

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package, for $z=0.201010$.

a. The comparisons are done at `r nrow(xi_2h)` unique values of $r$
a. Over the range tested 
($`r min(xi_2h$r)` < r < `r max(xi_2h$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(xi_2h$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(xi_2h$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$. 
a. `r testresult(xi_2h)`

The values of $\xi$ cover a very wide range, so a logarithmic $y$-axis is needed.
However, many points in the plot have $\xi = 0$.
In order to allow the use of a logarithmic axis, these points are suppressed.

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close. Note that we suppress negative values of $\\xi$ in this plot, so that we can use a log scale."}
ggplot(filter(xi_2h, ytest > 0.0, ytrue > 0.0)) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab(expression(r)) +
  ylab(expression(xi))
```

```{r fig.cap = "Relative residual comparison or our implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test. Note that points for which both the predicted and expected values are zero are suppressed."}
ggplot(filter(xi_2h, ytest > 0.0, ytrue > 0.0), 
       aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  #ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab(expression(r)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

### $\Sigma$ NFW

```{r}
local.frac.errormargin = 0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
# sig_nfw <- read_tsv("../validation-data/deltasigma_Sigma_nfwonly.out")
sig_nfw <- read_tsv("../validation-data/deltasigma_Sigma_nfwonly.out")
print(sig_nfw)
sig_nfw <- mutate(sig_nfw,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\Sigma_{NFW}(R)$ is the cluster surface mass density profile in the 1-halo regime, adopting an NFW model.

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package, for a cluster of $M_{200m}=3.199267137797384375e+14 M_\odot/h$ and $c=5$.

a. The comparisons are done at `r nrow(sig_nfw)` unique values of $R$.
a. Over the range tested 
($`r min(sig_nfw$r)` < R < `r max(sig_nfw$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(sig_nfw$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(sig_nfw$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(sig_nfw)`

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(sig_nfw) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("R [Mpc/h]") +
  ylab(expression(Sigma))
```

```{r fig.cap = "Relative residual comparison or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(sig_nfw, aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab("R [Mpc/h]") +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

### $\Sigma$ 2-halo

```{r}
local.frac.errormargin = 0.01
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
#sig_nfw <- read_tsv("../validation-data/deltasigma_Sigma_mm_2halo.out")
sig_nfw <- read_tsv("../validation-data/deltasigma_Sigma_mm_2halo.out")
# We cut off the range of R values at 30 Mpc. It would be better to do
# this by controlling the generation of the Sigma_mm_2halo.out file.
sig_nfw <- filter(sig_nfw, r <= 30)
sig_nfw <- mutate(sig_nfw,
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$\Sigma_{mm}(R)$ is the cluster surface mass density profile in the cluster 2-halo regime, adopting a non-linear matter power spectrum, and not accounting for cluster bias.

a. The comparisons are done at `r nrow(sig_nfw)` unique values of $R$.
a. Over the range tested 
($`r min(sig_nfw$r)` < R < `r max(sig_nfw$r)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(sig_nfw$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(sig_nfw$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(sig_nfw)`

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(sig_nfw) +
  geom_smallpoint(aes(x = r, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = r, y = ytest), color = testcolor, shape = testshape) +
  scale_x_log10() +
  scale_y_log10() +
  xlab("R [Mpc/h]") +
  ylab(expression(Sigma))
```

```{r fig.cap = "Relative residual comparison or our approximation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(sig_nfw, aes(x = r, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  scale_x_log10() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab("R [Mpc/h]") +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

### Halo bias $b(M_{200m}, z)$

```{r}
local.frac.errormargin =  0.001
local.abs.errormargin = abs.errormargin
```

```{r include = FALSE}
#bias <- read_tsv("../validation-data/deltasigma_bias.out")
bias <- read_tsv("../validation-data/deltasigma_bias.out")
bias$zbin <- as.factor(bias$z)
bias <- mutate(bias,
                 logmass  = log(M)/log(10),
                 resid = ytrue - ytest,
                 fracerr = resid/ytrue,
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```

$b(M_{200m}, z)$ is the halo bias at mass $M_{200m}$ and redshift $z$.

For this testing, we compare results from our implementation to the function evaluations from directly calling Tom McClintock's python/c package.

a. The comparisons are done at `r nlevels(bias$zbin)` different $z$ values,
and `r nunique(bias$logmass)` different $M_{200m}$ values.
a. Over the range tested 
($`r min(bias$logmass)` < Log M_{200m} < `r max(bias$logmass)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(bias$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(bias$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
or the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(bias)`

Each panel in the figure below corresponds to a $z$ bin;
the value of $z$ is specified in the panel top label.

```{r fig.cap = "Absolute implementation of our approximation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(bias) +
  geom_point(aes(x = logmass , y = ytrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = logmass , y = ytest), color = testcolor, shape = testshape) +
  scale_y_log10() +
  xlab("Log M200m") +
  ylab("Bias") +
  facet_wrap(zbin~.)
```

```{r, fig.cap = "Relative residual implementation of our approximation to the expected results. Blue points pass, and red points fail, the absolute residual test.", warning = FALSE}
ggplot(bias, aes(x = logmass, y = fracerr, color = within.abs)) +
  geom_point() +
  guides(color = "none") +
  scale_y_log10() +
  xlab("Log M200m") +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color = "blue") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  facet_wrap(zbin~ .)
```

## Mass Profile corrections

### Mis-centering offset

```{r}
local.frac.errormargin = frac.errormargin
local.abs.errormargin = abs.errormargin
```

```{r, include = FALSE}
miscent <- read_tsv("../validation-data/Rmis_gammadist_test.out")
miscent <- mutate(miscent,
                  resid = probtrue - probtest,
                  fracerr = resid/probtrue,
                  lo = pmin(-local.frac.errormargin * probtrue, -local.abs.errormargin),
                  hi = pmax(local.frac.errormargin * probtrue, local.abs.errormargin),
                  within.frac = (abs(fracerr) < local.frac.errormargin),
                  within.abs = (abs(resid) < local.abs.errormargin),
                  pass = within.frac | within.abs)
```

$P(R_{mis})$ is the probabilistic distribution of the mis-centering offset $R_{mis}$.
Our calculation is implemented in the file `roffset_t.hh`.

For this testing, we compare our function output to the evaluations from Yuanyuan Zhang in the DES Y1 centering analysis (Zhang et al. 2019).

a. The comparisons are done at `r nrow(miscent)` unique $R_{mis}$ values.
a. Over the range tested
$(`r min(miscent$rmis)` < R_{mis} < `r max(miscent$rmis)`)$,
the maximum magnitude of the absolute residual between the expected value
and our approximation is
$`r max(abs(miscent$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(miscent$fracerr))`$.
a. For this test, we require the relative residuals to be less than
$`r local.frac.errormargin`$,
or the absolute residuals to be less than
$`r local.abs.errormargin`$.
a. `r testresult(miscent)`

```{r fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(miscent) +
  geom_point(aes(x = rmis, y = probtrue), color = truecolor, shape = trueshape) +
  geom_point(aes(x = rmis, y = probtest), color = testcolor, shape = testshape) +
  xlab(expression(R_mis)) +
  ylab(expression(Probability))
```

```{r fig.cap = "Relative residual comparison or our implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test."}
ggplot(miscent, aes(x = rmis, y = fracerr, color = within.abs)) +
  geom_point() +
  guides(color = "none") +
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab(expression(R_mis)) +
  ylab("fractional residual") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")
```

## Lensing

### Critical density

```{r}
local.frac.errormargin =  0.0001
local.abs.errormargin = 1.0e-21
```

```{r include = FALSE}
sig_crit <- read_tsv("../validation-data/Sigma_crit_test.out")
sig_crit$zbin <- as.factor(sig_crit$zl)
sig_crit <- mutate(sig_crit,
                 resid = ytrue - ytest,
                 fracerr = resid/(ytrue+as.numeric(ytrue==0)),
                 lo = pmin(-local.frac.errormargin * ytrue, -local.abs.errormargin, 0),
                 hi = pmax(local.frac.errormargin * ytrue, local.abs.errormargin),
                 within.frac = (abs(fracerr) < local.frac.errormargin),
                 within.abs = (abs(resid) < local.abs.errormargin),
                 pass = within.frac | within.abs)
```


$\Sigma_{crit}(z_l, z_s)$ is the inverse of the critical surface density of a lens system, which changes with the redshift of the lens $z_l$ and the redshift of the source galaxy $z_s$. Our calculation is implemented in the file `sigma_crit_inverse_t.hh`. 

For this testing, we compare our function output to the evaluations using the astropy package with a flat cosmology $\Omega_m=0.3$, $\Omega_b=0.06$, $h_0=0.7$.

a. The comparisons are done at, `r nunique(sig_crit$zs)` unique $z_s$ values, with `r nlevels(sig_crit$zbin)` different $z_l$ values
a. Over the range tested 
 ($`r min(sig_crit$zs)` < zs < `r max(sig_crit$zs)`$),
the maximum magnitude of the absolute residual between the expected value
and our calculated value is
$`r max(abs(sig_crit$resid))`$;
the maximum magnitude of the fractional residual is
$`r max(abs(sig_crit$fracerr))`$.
a. For this test,
we require the relative differences to be 
less than $`r local.frac.errormargin`$,
**or** the absolute difference to be less than
$`r local.abs.errormargin`$.
a. `r testresult(sig_crit)`

Each panel in the figure below corresponds to a $\lambda_{c}$ bin;
the value of the $\lambda_{c}$ mass is specified in the panel top label.

```{r fig.asp=0.7, fig.cap = "Absolute comparison of our implementation to the expected results. Note that over-plotting may obscure the expected values, if the approximated values are very close."}
ggplot(sig_crit) +
  geom_smallpoint(aes(x = zs, y = ytrue), color = truecolor, shape = trueshape) +
  geom_smallpoint(aes(x = zs , y = ytest), color = testcolor, shape = testshape) +
  xlab("z_s") +
  ylab("Density [Mpc^(-3)]") +
  facet_wrap(vars(zbin))
```

```{r fig.asp=0.7, fig.cap = "Absolute residual comparison of our implementation to the expected results. Blue points pass, and red points fail, the relative residual test.", warning = FALSE}
ggplot(sig_crit, aes(x= zs, y = resid, color = within.frac)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.abs.errormargin, color = "blue") +
  geom_hline(yintercept = -1 * local.abs.errormargin, color = "blue") +
  ylim(-2 * local.abs.errormargin, 2 * local.abs.errormargin) +
  xlab("z_s") +
  facet_wrap(vars(zbin))
```

```{r fig.asp=0.7, fig.cap = "Relative residual comparison or our implementation to the expected results. Points in blue pass, and points in red fail, the absolute residual test. In this plot, the relative residual is forced to be 0 when both the true and test values are 0, as is the case when zs <= zl"}
ggplot(sig_crit, aes(x = zs, y = fracerr, color = within.abs)) +
  geom_smallpoint() +
  guides(color = "none") +
  geom_hline(yintercept = local.frac.errormargin, color="blue") +
  geom_hline(yintercept = -1 * local.frac.errormargin, color="blue")+
  ylim(-2 * local.frac.errormargin, 2 * local.frac.errormargin) +
  xlab("z_s") +
  facet_wrap(vars(zbin))
```

## Mass Concentration Relation
For this test, we are comparing the mass-concentration relation we get from the pipeline to the mass concentration relation from Colossus

## Session information

The following R package versions were used to prepare this document.

```{r}
sessionInfo()
```

